---
title: "Protein Signatures as Biomarkers for Cardiovascular Disease in HIV Patients"
author: "Mathanki Mehra"
date: "2026-03-23"
output: html_document
---

## 1. Installing Packages

```{r install-packages, eval=FALSE}
install.packages(c("tidyverse","readr","janitor","pheatmap","ggplot2"))
install.packages("BiocManager")
install.packages(c("igraph", "ggraph", "tidygraph"))
BiocManager::install(c("limma","clusterProfiler","org.Hs.eg.db","umap"))
BiocManager::install("ReactomePA")

```

## 2.Load all R libraries needed for data processing and visualization.
```{r load-libraries}
library(BiocManager)
library(dplyr)
library(tidyverse)
library(readr)
library(janitor)
library(limma)
library(pheatmap)
library(clusterProfiler)
library(org.Hs.eg.db)
library(ggplot2)
library(umap)
library(ReactomePA)
library(tibble)
library(knitr)
library(mitch)
library(stringr)
library(msigdbr)
library(eulerr)
library(beeswarm)
```

## 3. Load files

```{r load-files}
setwd("/mnt/data/mathanki/CVD_TMT")

plexA <- read_tsv("abundance_protein_MD_plexA_new.tsv")
plexB <- read_tsv("abundance_protein_MD_plexB_new.tsv")

plexA <- janitor::clean_names(plexA)
plexB <- janitor::clean_names(plexB)

```

## 4. Keep relevant columns

```{r keep-columns}
plexA <- dplyr::select(plexA, protein, gene,
                       x126, x127n, x127c, x128n, x128c,
                       x129n, x129c, x130n, x130c,
                       x131n, x131c, x132n)

plexB <- dplyr::select(plexB, protein, gene,
                       x126, x127n, x127c, x128n, x128c,
                       x129n, x129c, x130n, x130c,
                       x131n, x131c, x132n)
```

## 5. Rename TMT channels to sample IDs

```{r rename-channels}
annotationA <- c(
  x126="19191", x127n="34629", x127c="34862",
  x128n="40431", x128c="40475", x129n="35038",
  x129c="34866", x130n="41436", x130c="41383",
  x131n="40316", x131c="3746", x132n="102953"
)

annotationB <- c(
  x126="40797", x127n="4177", x127c="103376",
  x128n="47542", x128c="47442", x129n="42977",
  x129c="43081", x130n="113835", x130c="113856",
  x131n="19584", x131c="118117", x132n="118008"
)

colnames(plexA)[match(names(annotationA), colnames(plexA))] <- annotationA
colnames(plexB)[match(names(annotationB), colnames(plexB))] <- annotationB

```

## 6. Merge plexes

```{r merge-plexes}
merged <- full_join(plexA, plexB, by=c("protein","gene"))
head(merged)
```

## 7. Expression matrix for Merged plexes, Plex A and Plex B

```{r create-matrix}
expr_combined <- as.data.frame(merged)
protein_info_combined <- expr_combined[, c("protein", "gene")]
rownames(expr_combined) <- expr_combined$protein
expr_combined <- expr_combined[, !(colnames(expr_combined) %in% c("protein","gene"))]
expr_combined <- as.matrix(expr_combined)
storage.mode(expr_combined) <- "numeric"

expr_A <- as.data.frame(plexA)
protein_info_A <- expr_A[, c("protein", "gene")]
rownames(expr_A) <- expr_A$protein
expr_A <- expr_A[, !(colnames(expr_A) %in% c("protein","gene"))]
expr_A <- as.matrix(expr_A)
storage.mode(expr_A) <- "numeric"
head(expr_A)

expr_B <- as.data.frame(plexB)
protein_info_B <- expr_B[, c("protein", "gene")]
rownames(expr_B) <- expr_B$protein
expr_B <- expr_B[, !(colnames(expr_B) %in% c("protein","gene"))]
expr_B <- as.matrix(expr_B)
storage.mode(expr_B) <- "numeric"
head(expr_B)

```

## 8. Filter proteins (present in >=50% of samples)

```{r}
keep_combined <- rowSums(!is.na(expr_combined)) >= ncol(expr_combined)*0.5
expr_combined_filt <- expr_combined[keep_combined, ]

keep_A <- rowSums(!is.na(expr_A)) >= ncol(expr_A)*0.5
expr_A_filt <- expr_A[keep_A, ]


keep_B <- rowSums(!is.na(expr_B)) >= ncol(expr_B)*0.5
expr_B_filt <- expr_B[keep_B, ]

```


## 9. QC before normalization

```{r qc-pre-normalization}
qc_sample_combined <- data.frame(
  Sample = colnames(expr_combined_filt),
  Detected_Proteins = colSums(!is.na(expr_combined_filt)),
  Missing_Percent = colMeans(is.na(expr_combined_filt)) * 100,
  Median_Intensity = apply(expr_combined_filt, 2, median, na.rm = TRUE),
  SD_Intensity = apply(expr_combined_filt, 2, sd, na.rm = TRUE)
)

# Assign plex
plexA_samples <- c(
  "19191","34629","34862","40431","40475","35038",
  "34866","41436","41383","40316","3746","102953"
)

qc_sample_combined$Plex <- ifelse(qc_sample_combined$Sample %in% plexA_samples, "A", "B")

# QC summary by plex
qc_plex_combined <- qc_sample_combined %>%
  group_by(Plex) %>%
  summarise(
    N_Samples = n(),
    Mean_Detected_Proteins = mean(Detected_Proteins),
    Mean_Missing_Percent = mean(Missing_Percent),
    Mean_Median_Intensity = mean(Median_Intensity),
    .groups = "drop"
  )


kable(qc_sample_combined)
kable(qc_plex_combined)


# QC plots
qc_long_combined <- as.data.frame(expr_combined_filt) %>%
  rownames_to_column("Protein") %>%
  pivot_longer(-Protein, names_to = "Sample", values_to = "Intensity") %>%
  filter(is.finite(Intensity))

# Missingness plot
ggplot(qc_sample_combined,
       aes(x = reorder(Sample, Missing_Percent),
           y = Missing_Percent,
           fill = Plex)) +
  geom_col() +
  coord_flip() +
  theme_bw() +
  labs(title = "Missingness per Sample",
       x = "Sample", y = "% Missing")

# Detected proteins plot
ggplot(qc_sample_combined,
       aes(x = reorder(Sample, Detected_Proteins),
           y = Detected_Proteins,
           fill = Plex)) +
  geom_col() +
  coord_flip() +
  theme_bw() +
  labs(title = "Detected Proteins per Sample",
       x = "Sample", y = "Detected Proteins")

# Density plot
ggplot(qc_long_combined,
       aes(x = Intensity, color = Sample)) +
  geom_density(linewidth = 0.8) +
  theme_bw() +
  labs(title = "Pre-Normalization Intensity Distribution Graph",
       x = "Intensity", y = "Density")


# Plex-level comparison
ggplot(qc_plex_combined,
       aes(x = Plex, y = Mean_Detected_Proteins, fill = Plex)) +
  geom_col() +
  theme_bw() +
  labs(title = "Mean Detected Proteins by Plex",
       x = "Plex", y = "Mean Detected Proteins")

```

## 10. Quantile normalization and log2 transform

```{r}
expr_combined_norm <- normalizeBetweenArrays(expr_combined_filt, method = "quantile")
expr_combined_log <- log2(expr_combined_norm + 1)
boxplot(expr_combined_log, main = "Post-Normalization for both plexes", las = 2)

# Plex A
expr_A_norm <- normalizeBetweenArrays(expr_A_filt, method = "quantile")
expr_A_log <- log2(expr_A_norm + 1)
boxplot(expr_A_log, main = "Post-Normalization for Plex A", las = 2)

# plex B
expr_B_norm <- normalizeBetweenArrays(expr_B_filt, method = "quantile")
expr_B_log <- log2(expr_B_norm + 1)
boxplot(expr_B_log, main = "Post-Normalization for Plex B", las = 2)
```

## 11. QC after normalization

```{r qc-post-normalization}
qc_long_norm_combined <- as.data.frame(expr_combined_log) %>%
  rownames_to_column("Protein") %>%
  pivot_longer(-Protein, names_to = "Sample", values_to = "Intensity") %>%
  filter(is.finite(Intensity))

ggplot(qc_long_norm_combined,
       aes(x = Intensity, color = Sample)) +
  geom_density(linewidth = 0.8) +
  theme_bw() +
  labs(title = "Post-Normalization Intensity Distribution Graph",
       x = "Log2 Intensity",
       y = "Density")
cor_post_combined <- cor(expr_combined_log, use = "pairwise.complete.obs")

```

## 12. Impute missing values (median)

```{r}
# Combine Plexes
expr_combined_imp <- expr_combined_log
for(i in 1:ncol(expr_combined_imp)){
  expr_combined_imp[is.na(expr_combined_imp[,i]), i] <- 
    median(expr_combined_imp[,i], na.rm = TRUE)
}

# Plex A
expr_A_imp <- expr_A_log

for(i in 1:ncol(expr_A_imp)){
  expr_A_imp[is.na(expr_A_imp[,i]), i] <- 
    median(expr_A_imp[,i], na.rm = TRUE)
}

# Plex B
expr_B_imp <- expr_B_log

for(i in 1:ncol(expr_B_imp)){
  expr_B_imp[is.na(expr_B_imp[,i]), i] <- 
    median(expr_B_imp[,i], na.rm = TRUE)
}
sum(is.na(expr_combined_imp))
sum(is.na(expr_A_imp))
sum(is.na(expr_B_imp))

# Removing Zero Variance Proteins
expr_combined_imp <- expr_combined_imp[apply(expr_combined_imp, 1, var) != 0, ]
expr_A_imp <- expr_A_imp[apply(expr_A_imp, 1, var) != 0, ]
expr_B_imp <- expr_B_imp[apply(expr_B_imp, 1, var) != 0, ]
```

## Metadata Creation

```{r}

# METADATA CREATION (SINGLE SOURCE OF TRUTH)

# Create combined metadata from expression matrix
meta_combined <- data.frame(
  Sample = colnames(expr_combined_imp)
)

# Define Plex (based on known Plex A samples)

plexA_samples <- c(
  "19191","34629","34862","40431","40475","35038",
  "34866","41436","41383","40316","3746","102953"
)

meta_combined$Plex <- ifelse(meta_combined$Sample %in% plexA_samples,
                             "A", "B")

meta_combined$Plex <- factor(meta_combined$Plex)


# Define biological group (Case vs Control)


case_samples <- c(
  "19191","40797","34862","103376","40475","47442",
  "34866","43081","41383","113856","3746","118117"
)

meta_combined$Group <- ifelse(meta_combined$Sample %in% case_samples,
                              "Case", "Control")

meta_combined$Group <- factor(meta_combined$Group,
                              levels = c("Control","Case"))


# Final formatting


rownames(meta_combined) <- meta_combined$Sample


# Create Plex-specific metadata (CORRECT WAY)


meta_A <- meta_combined[meta_combined$Plex == "A", ]
meta_B <- meta_combined[meta_combined$Plex == "B", ]


# Align with expression matrices 

meta_A <- meta_A[colnames(expr_A_imp), ]
meta_B <- meta_B[colnames(expr_B_imp), ]

# Keep only needed columns


meta_A <- meta_A[, c("Sample","Group")]
meta_B <- meta_B[, c("Sample","Group")]

# SANITY CHECKS

# Alignment
all(rownames(meta_A) == colnames(expr_A_imp))
all(rownames(meta_B) == colnames(expr_B_imp))
all(rownames(meta_combined) == colnames(expr_combined_imp))

# Distribution checks
table(meta_combined$Plex, meta_combined$Group)
table(meta_A$Group)
table(meta_B$Group)
```

## PCA Before Batch Correction for combined plexes

```{r}
# PCA before batch correction
pca_pre <- prcomp(t(expr_combined_imp), center = TRUE, scale. = TRUE)

# Variance explained
percentVar_pre <- round(100 * (pca_pre$sdev^2 / sum(pca_pre$sdev^2)), 1)

# PCA dataframe
pca_df_pre <- data.frame(
  PC1 = pca_pre$x[,1],
  PC2 = pca_pre$x[,2],
  Sample = colnames(expr_combined_imp),
  Plex = meta_combined$Plex,
  Group = meta_combined$Group
)

# Plot
ggplot(pca_df_pre,
       aes(x = PC1, y = PC2,
           color = Plex,
           shape = Group)) +
  geom_point(size = 4) +
  geom_text(aes(label = Sample),
            vjust = -0.8,
            size = 3,
            show.legend = FALSE) +
  theme_bw() +
  labs(
    title = "PCA Before Batch Correction",
    x = paste0("PC1 (", percentVar_pre[1], "%)"),
    y = paste0("PC2 (", percentVar_pre[2], "%)")
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold")
  )
```

## Batch Correction

```{r}
expr_combined_corrected <- removeBatchEffect(
  expr_combined_imp,
  batch = meta_combined$Plex,
  design = model.matrix(~ Group, data = meta_combined)
)

# Quick check
head(expr_combined_corrected[,1:5])
# Dimensions unchanged
dim(expr_combined_corrected)
```

## PCA On Combined Plexes after removing Batch effect

```{r}
# PCA after batch correction
pca_post <- prcomp(t(expr_combined_corrected), center = TRUE, scale. = TRUE)

# Variance explained
percentVar_post <- round(100 * (pca_post$sdev^2 / sum(pca_post$sdev^2)), 1)

# PCA dataframe
pca_df_post <- data.frame(
  PC1 = pca_post$x[,1],
  PC2 = pca_post$x[,2],
  Sample = colnames(expr_combined_corrected),
  Plex = meta_combined$Plex,
  Group = meta_combined$Group
)

# Plot
ggplot(pca_df_post,
       aes(x = PC1, y = PC2,
           color = Group,
           shape = Plex)) +
  geom_point(size = 4) +
  geom_text(aes(label = Sample),
            vjust = -0.8,
            size = 3,
            show.legend = FALSE) +
  theme_bw() +
  labs(
    title = "PCA After Batch Correction",
    x = paste0("PC1 (", percentVar_post[1], "%)"),
    y = paste0("PC2 (", percentVar_post[2], "%)")
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold")
  )
```

```{r hierarchical-clustering-heatmap, fig.width=8, fig.height=7}

library(pheatmap)

# Sample correlation matrix
cor_matrix <- cor(expr_combined_corrected,
                  use = "pairwise.complete.obs")

# Sample annotations
annotation_df <- data.frame(
  Plex = meta_combined$Plex,
  Group = meta_combined$Group
)

rownames(annotation_df) <- colnames(expr_combined_corrected)

# Heatmap with hierarchical clustering
pheatmap(
  cor_matrix,
  annotation_col = annotation_df,
  annotation_row = annotation_df,
  clustering_method = "complete",
  main = "Hierarchical Clustering of Samples",
  fontsize = 10,
  border_color = NA
)
```



## Differential Expression Analysis
```{r}
# 1. PLEX A
design_A <- model.matrix(~ Group, data = meta_A)
fit_A <- lmFit(expr_A_imp, design_A)
fit_A <- eBayes(fit_A)
results_A <- topTable(
  fit_A,
  coef = "GroupCase",
  number = Inf,
  sort.by = "P"
)
results_A$Protein <- rownames(results_A)
# Add gene annotation
results_A <- merge(
  results_A,
  protein_info_A,
  by.x = "Protein",
  by.y = "protein",
  all.x = TRUE
)
# Add significance labels
results_A$Significance <- "Not Significant"
results_A$Significance[
  results_A$adj.P.Val < 0.05 & results_A$logFC > 1
] <- "Upregulated"
results_A$Significance[
  results_A$adj.P.Val < 0.05 & results_A$logFC < -1
] <- "Downregulated"
# View
head(results_A)

# 2. PLEX B
design_B <- model.matrix(~ Group, data = meta_B)
fit_B <- lmFit(expr_B_imp, design_B)
fit_B <- eBayes(fit_B)
results_B <- topTable(
  fit_B,
  coef = "GroupCase",
  number = Inf,
  sort.by = "P"
)
results_B$Protein <- rownames(results_B)
# Add gene annotation
results_B <- merge(
  results_B,
  protein_info_B,
  by.x = "Protein",
  by.y = "protein",
  all.x = TRUE
)
# Add significance labels
results_B$Significance <- "Not Significant"
results_B$Significance[
  results_B$adj.P.Val < 0.05 & results_B$logFC > 1
] <- "Upregulated"
results_B$Significance[
  results_B$adj.P.Val < 0.05 & results_B$logFC < -1
] <- "Downregulated"
# View
head(results_B)


# 3. Combined Plex
design_combined <- model.matrix(~ Plex + Group, data = meta_combined)
fit_combined <- lmFit(expr_combined_imp, design_combined)
fit_combined <- eBayes(fit_combined)
results_combined <- topTable(
  fit_combined,
  coef = "GroupCase",
  number = Inf,
  sort.by = "P"
)
results_combined$Protein <- rownames(results_combined)
# Add gene annotation
results_combined <- merge(
  results_combined,
  protein_info_combined,
  by.x = "Protein",
  by.y = "protein",
  all.x = TRUE
)
# Add significance labels
results_combined$Significance <- "Not Significant"

results_combined$Significance[
  results_combined$adj.P.Val < 0.05 & results_combined$logFC > 1
] <- "Upregulated"
results_combined$Significance[
  results_combined$adj.P.Val < 0.05 & results_combined$logFC < -1
] <- "Downregulated"
# View
head(results_combined)
```

## Overlap Of top 50 Up and down regulated genes for plex A and Plex B.

```{r}
# Plex A - Top 50 UP
top_A_up <- results_A %>%
  dplyr::arrange(P.Value) %>%
  dplyr::filter(logFC > 0) %>%
  dplyr::slice_head(n = 50)

# Plex A - Top 50 DOWN
top_A_down <- results_A %>%
  dplyr::arrange(P.Value) %>%
  dplyr::filter(logFC < 0) %>%
  dplyr::slice_head(n = 50)

# Plex B - Top 50 UP
top_B_up <- results_B %>%
  dplyr::arrange(P.Value) %>%
  dplyr::filter(logFC > 0) %>%
  dplyr::slice_head(n = 50)

# Plex B - Top 50 DOWN
top_B_down <- results_B %>%
  dplyr::arrange(P.Value) %>%
  dplyr::filter(logFC < 0) %>%
  dplyr::slice_head(n = 50)

A_up <- top_A_up$Protein
A_down <- top_A_down$Protein

B_up <- top_B_up$Protein
B_down <- top_B_down$Protein

shared_up <- intersect(A_up, B_up)
unique_A_up <- setdiff(A_up, B_up)
unique_B_up <- setdiff(B_up, A_up)

shared_down <- intersect(A_down, B_down)
unique_A_down <- setdiff(A_down, B_down)
unique_B_down <- setdiff(B_down, A_down)

cat("UP overlap:\n")
cat("Shared:", length(shared_up), "\n")
cat("Unique A:", length(unique_A_up), "\n")
cat("Unique B:", length(unique_B_up), "\n\n")

cat("DOWN overlap:\n")
cat("Shared:", length(shared_down), "\n")
cat("Unique A:", length(unique_A_down), "\n")
cat("Unique B:", length(unique_B_down), "\n")

fit_up <- euler(c(
  "Plex A" = length(A_up),
  "Plex B" = length(B_up),
  "Plex A&Plex B" = length(shared_up)
))

plot(fit_up,
     fills = c("#1f77b4", "#ff7f0e"),
     labels = TRUE,
     main = "Top 50 Upregulated Proteins Overlap")

fit_down <- euler(c(
  "Plex A" = length(A_down),
  "Plex B" = length(B_down),
  "Plex A&Plex B" = length(shared_down)
))

plot(fit_down,
     fills = c("#2ca02c", "#d62728"),
     labels = TRUE,
     main = "Top 50 Downregulated Proteins Overlap")
```


## Volcano plot

```{r volcano-combined-blue-white-red, fig.width=8, fig.height=6}
library(ggplot2)
library(ggrepel)
library(dplyr)

# Prepare volcano plot data
volcano_combined <- results_combined %>%
  mutate(
    negLog10P = -log10(P.Value),
    Label = ifelse(rank(P.Value, ties.method = "first") <= 10, gene, NA)
  )

# Volcano plot with blue-white-red scale
ggplot(volcano_combined,
       aes(x = logFC, y = negLog10P, fill = logFC)) +
  geom_point(
    shape = 21,
    color = "black",
    stroke = 0.2,
    alpha = 0.85,
    size = 2.5
  ) +
  geom_vline(
    xintercept = c(-1, 1),
    linetype = "dashed",
    linewidth = 0.5
  ) +
  geom_hline(
    yintercept = -log10(0.05),
    linetype = "dashed",
    linewidth = 0.5
  ) +
  geom_text_repel(
    aes(label = Label),
    size = 3,
    max.overlaps = 20,
    na.rm = TRUE
  ) +
  scale_fill_gradient2(
    low = "blue",
    mid = "white",
    high = "red",
    midpoint = 0,
    name = "log2 Fold Change"
  ) +
  theme_bw() +
  labs(
    title = "Differential Protein Abundance in Combined Plex Dataset",
    x = "log2 Fold Change",
    y = "-log10(P-value)"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold")
  )
```


## Top 50 up and down regulated proteins in combined plex.

```{r}
# Combined Plex - Top 50 UP
top_combined_up <- results_combined %>%
  dplyr::arrange(P.Value) %>%
  dplyr::filter(logFC > 0) %>%
  dplyr::slice_head(n = 50)

# Combined Plex - Top 50 DOWN
top_combined_down <- results_combined %>%
  dplyr::arrange(P.Value) %>%
  dplyr::filter(logFC < 0) %>%
  dplyr::slice_head(n = 50)

# View tables
kable(
  top_combined_up,
  caption = "Top 50 Upregulated Proteins in Combined Plex"
)

kable(
  top_combined_down,
  caption = "Top 50 Downregulated Proteins in Combined Plex"
)
```


## Combined Plex Pathway Enrichment Analysis
```{r}

# Combined Plex

rank_combined <- results_combined %>%
  dplyr::select(gene, t) %>%
  filter(!is.na(gene)) %>%
  group_by(gene) %>%
  summarise(t = mean(t), .groups = "drop") %>%
  arrange(desc(t))

# Get gene sets (Reactome)
msig_reactome <- msigdbr(
  species = "Homo sapiens",
  category = "C2",
  subcategory = "CP:REACTOME"
)

reactome_sets <- split(msig_reactome$gene_symbol,
                       msig_reactome$gs_name)

gene_list_combined <- rank_combined$t
names(gene_list_combined) <- rank_combined$gene

df_combined <- data.frame(logFC = gene_list_combined)

mitch_combined_reactome <- mitch_calc(
  df_combined,
  reactome_sets,
  minsetsize = 5
)

res_combined_reactome <- 
  mitch_combined_reactome$enrichment_result %>%
  arrange(p.adjustANOVA)

head(res_combined_reactome, 20)

unlink("mitchreport.html")
mitch_report(mitch_combined_reactome,"mitchreport.html")

```



## Figure
```{r reactome-dotplot, fig.width=10, fig.height=8}

library(ggplot2)
library(dplyr)

top_pathways <- res_combined_reactome %>%
  dplyr::select(set, setSize, pANOVA, s.dist, p.adjustANOVA) %>%
  arrange(p.adjustANOVA) %>%
  head(20)

top_pathways$set <- gsub("REACTOME_", "", top_pathways$set)
top_pathways$set <- gsub("_", " ", top_pathways$set)

ggplot(top_pathways,
       aes(x = s.dist,
           y = reorder(set, s.dist),
           size = setSize,
           color = p.adjustANOVA)) +
  geom_point(alpha = 0.9) +
  scale_color_gradient(
    low = "red",
    high = "blue",
    name = "Adjusted p-value"
  ) +
  theme_bw() +
  labs(
    title = "Top Enriched Reactome Pathways in Combined Plex",
    x = "Enrichment Score Distance",
    y = "Reactome Pathway",
    size = "Pathway Size"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text.y = element_text(size = 9),
    axis.title = element_text(face = "bold")
  )

```



## more figure
```{r pathway-network-plot, fig.width=10, fig.height=8}

library(igraph)
library(ggraph)
library(tidygraph)
library(dplyr)

# Select top pathways
network_df <- res_combined_reactome %>%
  arrange(p.adjustANOVA) %>%
  head(15)

# Clean names
network_df$set <- gsub("REACTOME_", "", network_df$set)
network_df$set <- gsub("_", " ", network_df$set)

# Create artificial pathway similarity edges
# (based on ranking adjacency for visualisation)

edges <- data.frame(
  from = network_df$set[-nrow(network_df)],
  to = network_df$set[-1]
)

# Create graph
graph <- graph_from_data_frame(edges, directed = FALSE)

# Add node attributes
V(graph)$pvalue <- network_df$p.adjustANOVA
V(graph)$size <- network_df$setSize

# Plot network
ggraph(graph, layout = "fr") +

  geom_edge_link(alpha = 0.5, colour = "grey70") +

  geom_node_point(aes(size = size,
                      color = pvalue)) +

  geom_node_text(aes(label = name),
                 repel = TRUE,
                 size = 3) +

  scale_color_gradient(
    low = "red",
    high = "blue",
    name = "Adjusted p-value"
  ) +

  theme_void() +

  labs(
    title = "Reactome Pathway Network",
    size = "Pathway Size"
  ) +

  theme(
    plot.title = element_text(hjust = 0.5,
                              face = "bold",
                              size = 14)
  )

```







## Session Info
```{r}
sessionInfo()
save.image("Analysis3.rdata")
```
